---
title       : 
subtitle    : 
author      : 
job         : 
framework   : revealjs
revealjs    : {theme: sky, transition: none, center: "false"} 
highlighter : highlight.js
hitheme     : github 
widgets     : [mathjax]
mode        : selfcontained 
url         : {lib: ./libraries}
knit        : slidify::knit2slides
assets:
  js:
    - "http://ajax.googleapis.com/ajax/libs/jqueryui/1.10.3/jquery-ui.min.js"
    - "http://bartaz.github.io/sandbox.js/jquery.highlight.js"
---



<style>

.reveal h4 {
    color: #380000;
    padding-bottom: 10px;
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal h3 {
    color: #380000;
    padding-bottom: 10px;
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal h2 {
    color: #380000;
    padding-bottom: 10px;
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal p {
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal body {
    font-family: 'Helvetica', 'Arial', sans-serif;
}


</style>




### formattable + slidify

<br>

[formattable](https://github.com/renkun-ken/formattable) is package by Kun Ren that makes formatting pretty tables in RMarkdown very easy!


<br>

[slidify](http://ramnathv.github.io/slidify/) is a package by Ramnath Vaidyanathan that makes making HTML slides in RMarkdown easy to make!

<br>

Here, I'm trying to use the two together...


---

### Example Table

<small>Data come from my [engsoccerdata](https://github.com/jalapic/engsoccerdata) package - these are summary stats for EPL teams 1992-2014 that played at least 400 games.</small>


```{r, echo=FALSE}
library(engsoccerdata)
library(dplyr)
specify_decimal <- function(x, k) format(round(x, k), nsmall=k)
mydf <- engsoccerdata2 %>% filter(Season>=1992  & tier==1) 
myt <- rbind(mydf %>% select(team=home, opp=visitor, gf=hgoal, ga=vgoal), 
      mydf %>% select(team=visitor, opp=home, gf=vgoal, ga=hgoal)) %>%
  group_by(team) %>%
  summarize(gp = n(), 
            wins=sum(gf>ga), 
            losses=sum(ga>gf), 
            draws=sum(gf==ga), 
            points=(wins*3)+draws,
            ppg = round((points/gp),2),
            winpct = specify_decimal((wins/gp),2)
            ) %>%
  arrange(desc(points))

myt400 <- myt %>% filter(gp>400)
myt400
```

---

#### Change format of the table in the following ways:

<div style='text-align: left; font-size: 75%;'>

<br>

&#10137;&nbsp;&nbsp; winpct is blue if the value is greater than 75%-quantile

<br>
<br>

&#10137;&nbsp;&nbsp; games played is bold when the value is equal to 848

<br>
<br>

&#10137;&nbsp;&nbsp; points and ppg is colorized by rank

</div>


---

<small>
```{r, echo=F}
library(formattable)
formattable(myt400, list(
  winpct = formatter("span", style = function(x)
    style(color = ifelse(x > quantile(x, 0.75), "mediumblue", NA))),
  gp = formatter("span", style = function(x)
    style("font-weight" = ifelse(x == 848, "bold", NA))),
 points = formatter("span", style = function(x, m = 1 - x/max(x) * 0.8, ms = round(1-m)) 
    style("display" = "block", "border-radius" = "4px", "background-color" = rgb(0, m, 0),
          "padding-right" = "4px", "color" = rgb(1,1,1))),
   ppg = formatter("span", style = function(x, m = 1-x/max(x), ms = round(1-m))
    style("display" = "block", "border-radius" = "4px", "padding-right" = "4px",
          "background-color" = rgb(m, m, 0), "color" = rgb(ms,ms,ms)))))

```
</small>


---

#### The code

```{r,eval=FALSE}
library(formattable)
formattable(myt400, list(
  winpct = formatter("span", style = function(x)
    style(color = ifelse(x > quantile(x, 0.75), "mediumblue", NA))),
  gp = formatter("span", style = function(x)
    style("font-weight" = ifelse(x == 848, "bold", NA))),
 points = formatter("span", style = function(x, m = 1 - x/max(x) * 0.8, ms = round(1-m)) 
    style("display" = "block", "border-radius" = "4px", "background-color" = rgb(0, m, 0),
          "padding-right" = "4px", "color" = rgb(1,1,1))),
   ppg = formatter("span", style = function(x, m = 1-x/max(x), ms = round(1-m))
    style("display" = "block", "border-radius" = "4px", "padding-right" = "4px",
          "background-color" = rgb(m, m, 0), "color" = rgb(ms,ms,ms)))))

```

<br>
<small>Also, I wrapped the R code chunk in `<small>` tags</small>