---
title       : 
subtitle    : 
author      : 
job         : 
framework   : revealjs
revealjs    : {theme: sky, transition: none, center: "false"} 
highlighter : highlight.js
hitheme     : github 
widgets     : [mathjax]
mode        : selfcontained 
url         : {lib: ./libraries}
knit        : slidify::knit2slides
assets:
  js:
    - "http://ajax.googleapis.com/ajax/libs/jqueryui/1.10.3/jquery-ui.min.js"
    - "http://bartaz.github.io/sandbox.js/jquery.highlight.js"
---



<style>

.reveal h4 {
    color: #380000;
    padding-bottom: 10px;
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal h3 {
    color: #380000;
    padding-bottom: 10px;
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal h2 {
    color: #380000;
    padding-bottom: 10px;
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal p {
    font-family: 'Helvetica', 'Arial', sans-serif;
}

.reveal body {
    font-family: 'Helvetica', 'Arial', sans-serif;
}



</style>


## Making Tables in Slidify



--- 

This is the dataframe that I'm going to turn into a series of nice looking tables:

```{r}
df <- data.frame(State=c("Alabama", "Alaska", "Arizona", "Arkansas"),
           Flower=c("Camellia", "Forget-me-not", "Saguaro cactus blossom", "Apple blossom"),
           Amphibian=c("Red Hills Salamander", "none", "Arizona Tree Frog", "none"),
           Bird=c("Northern Flicker", "Willow ptarmigan", "Cactus wren", "Northern mockingbird")
           )
df
```

<br>
<small>It doesn't neatly fit across the screen</small>

---

The HTML code for writing this table would be as follows:

```
<table id="tabletype" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


```

<br>

<small>`tabletype` refers to the series of instructions in the css style sheet belonging to that table type.</small>

---

For instance, here is the style sheet for a very simple grid table type:

```
/*GRIDTABLE STYLE*/
table#gridtable {font-family: 'Helvetica', 'Arial', sans-serif; font-size:16px; color:#333333; border-width: 1px; border-color: #666666; border-collapse:collapse;}
table#gridtable th {font-size:18px;border-width: 1px; padding: 8px; border-style: solid;border-color: #666666; background-color: #dedede;}
table#gridtable td {border-width: 1px; padding: 8px; border-style: solid; border-color: #666666; background-color: #ffffff;}
```

<br>
<small>The result of this is on the next slide</small>


---

#### Simple Grid Table type


<br>


<table id="gridtable" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>

---  

The following slides give you some example tables styles that can be written into the custom css and referred to in slidify presentations.  They are all available at [this github page](https://github.com/psych-git/psych-slides/tree/master/revealjs/assets/css).

<br>

The first set of table examples require you to write out the table in the long HTML format.  This is fine for tiny tables like the example here, but would be super annoying for any large table.   There are options for turning R output into HTML style code, but I haven't worked out what might be the best solution just yet.  In the last few slides I introduce a little function I wrote to turn a data.frame straight into a nice looking table in one line !

<br>
All of the following are created by simply modifying the custom css.

---

#### Another very Simple style

<br>


<table id="simple" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---

#### Zebra Style with colored lines 

<br>

<table id="zebra1" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---

#### Zebra Style with solid lines and header filled

<br>

<table id="zebra" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---

#### Box - Rows change color when hovered over


<br>


<table id="box" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>

---

#### Another box style

<br>


<table id="onecol" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


--- 

#### Minimal - Text changes weight when hovered over


<br>


<table id="minimal" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---

#### Another Simple Hover Table


<br>


<table id="hoverTable" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---


#### Clean, modern looking hover table

<br>

<table id="flattable" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>

---



#### A green version of the previous example

<br>

<table id="flattable3" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---

#### Another table

<br>

<table id="another" style="width:100%">
  <tr>
    <th>State</th>
    <th>Flower</th>
    <th>Amphibian</th>
    <th>Bird</th>
  </tr>
  <tr>
    <td data-th="State">Alabama</td>
    <td data-th="Flower">Camellia</td>
    <td data-th="Amphibian">Red Hills Salamander</td>
    <td data-th="Bird">Northern Flicker</td>
  </tr>
  <tr>
    <td data-th="State">Alaska</td>
    <td data-th="Flower">Forget-me-not</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Willow ptarmigan</td>
  </tr>
  <tr>
    <td data-th="State">Arizona</td>
    <td data-th="Flower">Saguaro cactus blossom</td>
    <td data-th="Amphibian">Arizona Tree Frog</td>
    <td data-th="Bird">Cactus wren</td>
  </tr>
  <tr>
    <td data-th="State">Arkansas</td>
    <td data-th="Flower">Apple blossom</td>
    <td data-th="Amphibian">none</td>
    <td data-th="Bird">Northern mockingbird</td>
  </tr>
</table>


---

<br>
Making these tables is great, but nobody wants to write out a bunch of HTML code to make a table.  It's easy to make an error, it's boring and it's annoying.  

<br>

I have a temporary solution - I have written a little function that takes two arguments. <br><br> 1) the <font color="red">`name of the dataframe`</font> that I want to turn into a pretty looking table on my slide <br> <br> 2) the <font color="red">`tabletype`</font> I want it to end up as - i.e. the css format guide for tables that I've predefined in my css.



---

<br>

This is the function - it's ugly. If anybody wants to help me improve it, then great.
<br>
<br>

```{r}
tablecode <- function(df, tabletype="gridtable"){
tmp <- noquote(apply(df, 1, function(x) paste0("<td data-th='", sprintf("%s",colnames(df)), "'>", x, "</td>", collapse=" ")))

cat(
c(
noquote(paste0("<table id='", tabletype,"' style='width:100%'> ")),    
noquote(paste0("<tr>", paste0("<th>", sprintf("%s",colnames(df)), "</th>", collapse=" "), "</tr>")),
noquote(unlist(lapply(tmp, function(x) paste0("<tr>", x, "</tr>")))),
noquote("</table>")
)
)

}
```

---

<br>

The default `tabletype` is <font color="red">gridtable</font>
<br>
<br>The other types you can use are:  <font color="red">zebra, zebra1, box, minimal, onecol, simple, hoverTable, flattable, flattable3, another</font>.
<br><br>

You have to have added the css for these into your `custom.css` file in your `assets/css` folder.

---

<br>
Let's try this using a different table.
<br>
<br>

```{r}

paintings <- data.frame(
Painting = c("Nafea Faa Ipoipo", "The Card Players", "Le Reve", "Three Studies of Lucian Freud", "No. 5, 1948"),
Artist = c("Paul Gauguin", "Paul Cezanne", "Pablo Picasso", "Francis Bacon", "Jackson Pollock"),
Year = c(1892, 1892, 1932, 1969, 1948),
Price = c("$300M", "$259M", "$155M", "$142M", "$140M")
)

paintings

```

---

The function is very simple:
<br>

```{r}
tablecode(paintings, tabletype="onecol")
```

<br>

This output can simply be copied and pasted into the RMarkdown document between two slide separators to produce the desired table...

---


I copied and pasted the output of `tablecode(paintings)` into my RMarkdown document and it works!

<br>

<table id='onecol' style='width:100%'>  <tr><th>Painting</th> <th>Artist</th> <th>Year</th> <th>Price</th></tr> <tr><td data-th='Painting'>Nafea Faa Ipoipo</td> <td data-th='Artist'>Paul Gauguin</td> <td data-th='Year'>1892</td> <td data-th='Price'>$300M</td></tr> <tr><td data-th='Painting'>The Card Players</td> <td data-th='Artist'>Paul Cezanne</td> <td data-th='Year'>1892</td> <td data-th='Price'>$259M</td></tr> <tr><td data-th='Painting'>Le Reve</td> <td data-th='Artist'>Pablo Picasso</td> <td data-th='Year'>1932</td> <td data-th='Price'>$155M</td></tr> <tr><td data-th='Painting'>Three Studies of Lucian Freud</td> <td data-th='Artist'>Francis Bacon</td> <td data-th='Year'>1969</td> <td data-th='Price'>$142M</td></tr> <tr><td data-th='Painting'>No. 5, 1948</td> <td data-th='Artist'>Jackson Pollock</td> <td data-th='Year'>1948</td> <td data-th='Price'>$140M</td></tr> </table>

---

An even cooler solution is to add `results='asis'` to our R chunk options.  The below is what I would write at the top of my R chunk to produce the desired table:

<br>

```
{r, results='asis', echo=FALSE}
tablecode(paintings, tabletype="minimal")

```

<br>

The next slide shows the output of this simple one line piece of code ...

---

```{r, results='asis', echo=FALSE}
tablecode(paintings, tabletype="minimal")

```

<br>

---

<p style='font-size=12px'>
<font color = 'red'>It worked !!!   </font>We can now write really nice looking tables with one function without having to write out HTML tables in longhand.  

<br>
<br>

Just play around with all the settings in the different table templates in the custom.css file to get your table looking the way you want it. You can even add a new 'id' to the table to make it distinct.

<br><br>

If I get some  time, I may even work on extending this function to get it to add inline CSS to the RMarkdown document, thereby allowing the user to on the fly change cell heights, colors, padding, margins, font sizes etc. etc.  without having to go and change the custom.css file....   Anyone want to help then give me an email &#9786;
</p>


---

#### Adding inline CSS to the RMarkdown document

<br>

The next table is produced by simply adding in the following code after the slide divider in our RMarkdown document.

<br>

```
<style>
table#another1 {border-collapse: collapse; border: 1px solid #900000; 
background: #FFCCCC; font-family: 'Helvetica', 'Arial', sans-serif; font-size:16px; color: #003333; }
table#another1 td, th { padding: 3px 3px .75em 3px; line-height: 1.3em; }
table#another1 th {text-shadow: none; font-size:18px; background: #CC0000; color: white; font-weight: 200; text-align: left; padding-right: .1em;  padding-left: .2em; vertical-align: middle; }
</style>
```
<br>

and referring to it like this:

```
{r, results='asis', echo=FALSE}
tablecode(paintings, tabletype="another1")

```

---

<style>
table#another1 {border-collapse: collapse; border: 1px solid #900000; 
background: #FFCCCC; font-family: 'Helvetica', 'Arial', sans-serif; font-size:16px; color: #003333; }
table#another1 td, th { padding: 3px 3px .75em 3px; line-height: 1.3em; }
table#another1 th {text-shadow: none; font-size:18px; background: #CC0000; color: white; font-weight: 200; text-align: left; padding-right: .1em;  padding-left: .2em; vertical-align: middle; }
</style>


```{r, results='asis', echo=FALSE}
tablecode(paintings, tabletype="another1")

```

<br>

